Learning constraints in spreadsheets and tabular data

### 2016-09-23 Paper/ECAI/tabular.tex
Updated plots, plot references (and block detection)

\paragraph{Experimental setup}
\label{sec:evalualtion:method}
All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once (using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g. merge split-up tables because of None values, or remove headers of all-textual tables).
\samuel{Should we not say just with the visual tool?}
 \\ % split of weird sentence
Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) unless a spreadsheet contains ambiguous \textit{None} values.
For those ambiguous cases the blocks are specified manually. \samuel{I have yet to make this more precise}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
  \caption{Loglog plots of the experiments with synthetic (random) data. In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plots~$c$ and~$d$ the number of blocks is increased.}
  \label{fig:runtime_analysis}
\end{figure}

@@

Typically, we expect the number of generated subassignments to be much smaller in real world spreadsheets and, therefore, the running time to be less sensitive to the number of vectors.
This is because in phase~1 entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths for the product constraint.
% \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
\samuel{=> benchmark}

@@

As discussed in \samuel{???}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
Our experiments satisfy the assumption that the large majority of possible subassignments do not satisfy most constraints.
Therefore, the results confirm that our implementation is able to severely limit the impact of the length of vectors.

@@

\paragraph{Effect of block size}
To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are split into varying number of blocks (up to ~$32$).
Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
As the blocks become smaller, the total running time first decreases before increasing again.
Looking at the running times for aggregate constraints and non-aggregate constraints  separately reveals that they are affected differently by the block size.
For aggregate constraints (\samuel{xxx line}) the running time decreases with the block size, while for non-aggregate constraints (\samuel{yyy line}) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes very small.

@@

\paragraph{Q1: Effectiveness}
Our first questions, Q1, is about the inability to find intended constraints due to excessive runtime. As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.  
The average runtime across all spreadsheets is $0.13s$.
Across categories the average running times fluctuate in the range between $0.08s$ to $0.20s$.
While most spreadsheets can be processed very quickly, running times can go up to around $0.85s$ for others.
In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets∞m.

>>>>

\paragraph{Experimental setup}
\label{sec:evalualtion:method}
All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once (using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g. merge split-up tables because of None values, or remove headers of all-textual tables).
\samuel{Should we not say just with the visual tool?}
Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) except for spreadsheets that contain \textit{None} values or (partially) numeric headers, in which case the blocks are specified manually.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
  \caption{Loglog plots of the experiments with synthetic (random) data. In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plot~$c$ the number of blocks is increased (the block size is decreased).}
  \label{fig:runtime_analysis}
\end{figure}

@@

Typically, we expect the number of generated subassignments to be much smaller in real world spreadsheets and, therefore, the running time to be less sensitive to the number of vectors.
This is because in phase~1 entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths for the product constraint.
% \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}

@@

As discussed in Section~\ref{sec:which_cons}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
Our experiments satisfy the assumption that the large majority of possible subassignments do not satisfy most constraints.
Therefore, the results confirm that our implementation is able to severely limit the impact of the length of vectors.

@@

\paragraph{Effect of block size}
To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are split into varying number of blocks (up to ~$32$).
Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
As the blocks become smaller, the total running time first decreases before increasing again.
Looking at the running times for aggregate constraints and non-aggregate constraints  separately reveals that they are affected differently by the block size.
For aggregate constraints (green line) the running time decreases with the block size, while for non-aggregate constraints (red line) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes small.

@@

\paragraph{Q1: Effectiveness}
Our first questions, Q1, is about the inability to find intended constraints due to excessive runtime. As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.  
The average runtime across all spreadsheets is $0.13s$.
Across categories the average running times fluctuate in the range between $0.08s$ to $0.20s$.
While most spreadsheets can be processed very quickly, running times can go up to around $0.85s$ for others.
In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.
